---
title: "WUP_04"
author: "Chris Gonzalez"
date: "Summer 2023"
output:
  html_document:
    df_print: paged
---

```{r include=FALSE,echo=FALSE}
require(knitr)
require(tidyverse)
require(tigerstats)
thedata<-read.csv(file="GSScsv.csv")
```
```{r eval=FALSE}
thedata <- thedata %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

The data is in the file GSScsv.csv and came from the 2018 General Social Survey.  The question to answer is to resolve this hypothesis.  
$H_0$ There is no difference between the average score on the WORDSUM test and level of schooling DEGREE  
$H_a$ There is a significant difference in average score on the WORDSUM test between at least two levels of DEGREE  

# Methods

<!--Mention that you intend to use chatGPT to the greatest extent possible. Give credit to chatGPT when you use it to create code in your report. Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

Initially, our objective was to analyze the available data to assess its composition. We replicated the methodology utilized in write-up 3, modifying the character variables to factor variables. Through this process, we identified that we possessed numerical data represented by the variable "WORDSUM," while the variable "degree" was transformed into a factor variable. Consequently, we encountered a challenge known as a numerical ~ categorical problem.


# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->



## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->


```{r}
ggplot(thedata, aes(x = DEGREE, y = WORDSUM)) +
  geom_boxplot(fill = "lightgreen", color = "darkgreen", outlier.shape = NA) +
  theme_minimal() +
  labs(x = "DEGREE", y = "WORDSUM", title = "Fancy Box Plot") +
  theme(
    plot.title = element_text(hjust = 0.5, size = 16)
  )
```

The analysis of the dataset suggests that there is no overlap between individuals who have completed high school and those who have attended junior college or graduate school. This finding indicates a low likelihood of significant differences in educational background between these groups. Additionally, there is no overlap between individuals who have completed graduate school and those with a high school education, further supporting the notion of distinct educational attainment levels.It is noteworthy that the distinguishing factors in the dataset are limited to individuals with a high school education compared to those who have completed graduate studies. These differences highlight the significance of educational levels achieved, implying that these two categories are the primary drivers of differentiation within the dataset.



### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
favstats(WORDSUM~DEGREE,data=thedata)
```
The analysis of the data indicates that the average scores on the WORDSUM test for graduates were slightly above 10. Notably, there exists a statistically significant difference between the average scores observed in graduate school and junior college. These findings provide evidence that the educational experiences offered by graduate school and junior college are distinct from each other, as reflected by the disparities in average test scores.


## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
model1 <- aov(WORDSUM~DEGREE,data=thedata)
summary.aov(model1)
```
The initial analysis reveals a significant difference in educational attainment across all five degrees. However, it is important to note that further testing and analysis are required to delve deeper into the data and obtain a more comprehensive understanding of the observed differences. Additional tests can be conducted to explore the specific nature and magnitude of the disparities between the educational levels. These subsequent analyses will facilitate a more nuanced and comprehensive examination of the data, allowing for a deeper dissection of the observed patterns and relationships.

# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->
